Modelling fine-scale dive behavior with hidden Markov models

### 2020-07-08 Report/Sections/appendix.tex
Commit before changing STFT to Fourier Transform

>>>>

\section{Results regarding CarHMMs and the Ornstein-Uhlenbeck process}

\subsection{Continuous Time models}

Although the CarHMM can incorporate auto-correlation into the structure of an HMM, they break down when observations are taken at irregular time intervals since the probability transition matrix $\Gamma$ would then have to depend upon the time step between observations. A common solution to this issue is to use a continuous-time method such as the one introduced by Michelot et al \citep{Michelot:2019}, which models the movement of an animal as an Ornstein-Uhlenbeck process with parameters that depend upon the underlying behavioural state of the animal. Namely, a state-switching Ornstein-Uhlenbeck process $Y$ is the solution to the following stochastic differential equation:
%
\begin{equation}
    \label{eqn:SDE}
    dY_t = \beta^{(X_t)}(\gamma^{(X_t)} - Y_t)dt + \omega^{(X_t)} dW_t
\end{equation}
%
where $X_t$ follows some stochastic process which defines the hidden behaviour of the animal at time $t$, $\beta^{(X_t)}$ relates to rate at which the process returns to its mean value, $\gamma^{(X_t)}$ is the long-term mean value of the process, $\omega^{(X_t)}$ is related to short-term variance, and $W$ is a Wiener process. One notable difference between the state-switching OU process and the CarHMM is that the time index $t \in \mathbb{R}$ exists in continuous time and is not necessarily an integer. If the behavioral state $X_t$ is known and does not change between observations, the solution to (eqn \ref{eqn:SDE}) is known to be the following \citep{Michelot:2019}:
%
\begin{align}
Y_{t+\Delta t} | X_{t} \sim \mathcal{N}\left((1-e^{-\beta^{(X_t)}\Delta t})\gamma^{(X_t)} + e^{-\beta^{(X_t)}\Delta t} Y_t,\quad \frac{\omega^{(X_t)^2}}{2\beta^{(X_t)}} (1-e^{-2\beta^{(X_t)}\Delta t})\right)
\label{eqn:OU_sol}
\end{align}
%
where $\Delta t$ is the time difference between any two observations $Y_t$ and $Y_{t+\Delta t}$. Continuous time models can be built up from arbitrarily complex stochastic differential equations and allow for uneven step lengths in the observations sequence $Y$. However, most continuous time models require MCMC algorithms to perform inference and as a result are not easily incorporated into the HHMM structure.

\subsection{Equivalence of CarHMM and OU process}

Although most continuous time models are difficult to incorporate into an HHMM, we show in (thrm 1) that under certain conditions, the CarHMM is equivalent to a state-switching Ornstein-Uhlenbeck process shown in (eqn \ref{eqn:OU_sol}). This gives new interpretation to the learned parameters of the CarHMM in the context of a continuous-time model.

\begin{theorem}{Theorem 1.}{}%
If:
\begin{enumerate}
    \item The hidden behavioural process $X$ from (Eq. \ref{eqn:SDE}) follows a Markov chain with $N$ possible states and transitions occur at equi-spaced time stamps $\left(\Delta t, \ldots, (T-1)\Delta t\right)$, and:
    \item Observations of the SDE from (Eq. \ref{eqn:SDE}) are taken at times $\left(0, \Delta t, \ldots, (T-1)\Delta t\right)$,
\end{enumerate}
then the observations $y$ are equivalent to the output of a conditionally auto-regressive hidden Markov model with normal emission distributions and parameters $\Theta = (\theta^{(1)}, \ldots, \theta^{(N)}); \enspace \theta^{(i)} = \{\mu^{(i)},\sigma^{(i)},\phi^{(i)}\}$, where:

\begin{align}
\mu^{(i)} = \gamma^{(i)}, \qquad \sigma^{(i)} = \sqrt{\frac{\omega^{(i)^2}}{2\beta^{(i)}} (1-e^{-2\beta^{(i)}\Delta t})}, \qquad \phi^{(i)} = e^{-\beta^{(i)}\Delta t} \label{eqn:CarHMM_to_OU}
\end{align}

\end{theorem}

\begin{proof}{Proof}{}%
Combining equation (\ref{eqn:CarHMM_to_OU}) with equation (\ref{eqn:OU_sol}) gives:
%
\begin{align*}
\left(Y_{s \Delta t} | X_{(s-1)\Delta t} = i \right) \sim \mathcal{N}\left((1-\phi^{(i)}) \mu^{(i)} + \phi^{(i)} Y_{(s-1) \Delta t}, \enspace \sigma^{2(i)} \right),\\
s = 1, \ldots, T-1
\end{align*}
%
If $X_t$ follows a Markov chain with transitions at the observation times, then the behavioural state $X_t$ does not change between observations and we can re-index $Y_{(s-1) \Delta t} = Y'_s$ and $X_{(s-2)\Delta t: (s-1) \Delta t} = X'_s$ for $s = 2,\ldots T$, yielding the desired result:

\begin{align*}
\left(Y'_s| X'_s = i \right) \sim \mathcal{N}\left((1-\phi^{(i)}) \mu^{(i)} + \phi^{(i)} Y'_{s-1}, \enspace \sigma^{2(i)} \right)\\
s = 2, \ldots, T
\end{align*}
%
$X'$ follows a Markov chain, and the distribution of $(Y'_s|X'_s,Y'_{s+1})$ is consistent with that of a CarHMM with normal emission distributions.
\end{proof}

### 2020-07-08 Report/Sections/intro.tex
Commit before changing STFT to Fourier Transform

One of the most prevalent techniques in the literature of late is the hidden Markov model (HMM), where observations depend upon the state of an associated unobserved behavioral process following Markovian dynamics \citep{Patterson:2017}. Importantly, under the traditional HMM model, subsequent observations are assumed to be independent from one another conditioned on the underlying behavioral process. However, this assumption is often violated in the real world processes, especially when observations are taken a high frequencies. For example, the location of an animal at a given time is highly correlated with the location of that animal one second later. Several publications have dealt with this issue in the past, including the hidden movement Markov model (HMMM) by Whoriskey et al \citep{Whoriskey:2016} and the conditionally auto-regressive hidden Markov model (CarHMM) by Lawer et al. \citep{Lawler:2019}. The CarHMM in particular explicitly models auto-correlation into an HMM while maintaining the structure needed to run the forward algorithm. It also only adds one additional parameter per possible hidden state

@@

@@

One way to deal with auto-correlation in fine-scale behavioral processes is to use the CarHMM, or \textit{conditionally auto-regressive hidden Markov model}, introduced by Lawler et al \citep{Lawler:2019}. .....  The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to run the forward algorithm. It also only adds one additional parameter per possible hidden state
$(\phi^{(i)})$. 
\end{quote}

>>>>

One of the most prevalent techniques in the literature of late is the hidden Markov model (HMM), where observations depend upon the state of an associated unobserved behavioral process following Markovian dynamics \citep{Patterson:2017}. Importantly, under the traditional HMM model, subsequent observations are assumed to be independent from one another conditioned on the underlying behavioral process. However, this assumption is often violated in the real world processes, especially when observations are taken a high frequencies. For example, the location of an animal at a given time is highly correlated with the location of that animal one second later. Several publications have dealt with this issue in the past, including the hidden movement Markov model (HMMM) \citep{Whoriskey:2016} and the conditionally auto-regressive hidden Markov model (CarHMM) \citep{Lawler:2019}. The CarHMM in particular explicitly models auto-correlation into an HMM while maintaining the structure needed to run the forward algorithm. It also only adds one additional parameter per possible hidden state.

@@

For periodic behavior in particular, One way to avoid the use of continuous time models is to use signal processing techniques such as the Fourier transform on the raw data. The advantages of using Fourier analysis within an HMM has been recently demonstrated in the context of describing daily behavioral cycles of marine mammals \citep{Heerah:2017}. In addition, Fourier analysis has previously been used in the field of animal movement to explain animal behavior \citep{Fehlmann:2017} and specifically fluking \citep{Shorter:2017} from accelerometer data. Thus, incorporating Fourier analysis of accelerometer data within the structure of an HMM appeared a promising simple approach to account for additional correlation in data that is cyclical in nature.

@@

One way to deal with auto-correlation in fine-scale behavioral processes is to use the CarHMM, or \textit{conditionally auto-regressive hidden Markov model}, introduced by \citep{Lawler:2019}. .....  The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to run the forward algorithm. It also only adds one additional parameter per possible hidden state
$(\phi^{(i)})$. 
\end{quote}

### 2020-07-08 Report/Sections/models.tex
Commit before changing STFT to Fourier Transform

%
We denote the conditional distribution of $Y_t$ given $Y_{t-1}= y_{t-1}$ and $ X_t=i$ as $f^{(i)}( \cdot | y_{t-1})$ or as
$f^{(i)}( \cdot | y_{t-1}; \theta^{(i)})$.
As an example, we could assume that this conditional distribution is normal with parameters $\theta^{(i)} = \{\mu^{(i)},\sigma^{2(i)},\phi^{(i)}\}$, with:
%
\[
\mathbb{E}(Y_t|Y_{t-1} = y_{t-1},X_t=i) = \phi^{(i)} ~ y_{t-1} ~+ ~(1-\phi^{(i)})  ~\mu^{(i)}
\]
and
\[
\mathbb{V}(Y_t| Y_{t-1} =y_{t-1}, X_t = i) = \sigma^{2(i)}.
\]
%
{\bf{Ah - now I see the issue about changing states.  It would be great to have something about the distribution of $Y_t$ given $X_t=i, X_{t-1}=j$. You'd want a different conditional distribution, depending on if $i=j$ or not.
}}
%
The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to write the likelihood using the forward algorithm. In our normal model formulation, we have added only one additional parameter, $\phi^{(i)}$,  per possible hidden state.

The likelihood for the CarHMM can be easily calculated using the forward algorithm.  As previously, let $y$ be the vector of observed emissions.  Then
\begin{equation}
\calL_{\text{CarHMM}}(y;\Theta,\Gamma) = \delta \prod_{t=2}^T \Gamma P(y_t;\Theta) \mathbf{1}_N
\label{CarHMM_likelihood}
\end{equation}
where
%
$P(y_t;\Theta)$ is an $N \times N$ diagonal matrix with $ii$th entry equal to $f^{(i)}(y_t|y_{t-1}; \theta^{(i)})$.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The hierarchical hidden Markov model (HHMM)}

A hierarchical hidden Markov model contains two process levels, a coarse-scale process and a fine-scale process.  The coarse-scale process is a hidden Markov model as defined in the previous subsection,  with  $X_1\ldots, X_T$ the unobserved Markov chain of states and $Y_1,\ldots, Y_T$ the corresponding observed responses.  The possible states are $1,\ldots, N$.   
%

For each value of $t$, the state $X_t$ emits not only $Y_t$ but also  a sequence of unobserved states, 
$X_t^* \equiv (X_{t,1}^*,\ldots, X_{t,T_t^*})$ and a sequence of observed emissions, 
$Y_t^* \equiv (Y_{t,1}^*,\ldots, Y_{t,T_t^*})$. 
These form our fine-scale process.
Conditional on the coarse-scale hidden state $X_t$, the joint fine-scale process,  $X_t^*$, $Y_t^*$, is a hidden Markov model with parameters depending on the value of $X_t$.  Specifically, given that $X_t=i$,   $X_t^*$ is a Markov chain with states $1,\ldots, N_t^*$ with 
%
\hfill\break{\bf{NH:  Do we write $1^*$? That's strange.}}
\hfill\break
%
 distribution of $X_t^*$, conditional on $X_t=i$,  given by an $N^*_t \times N^*_t$ transition probability matrix $\Gamma^{*(i)}$ and initial probability, denoted by the $N_t^*$-vector $\delta^{*(i)}$, which we take equal to the stationary distribution of the chain.     For simplicity, we  take $N_t^* \equiv N^*$ although this is not necessary.
%
The distribution of $Y_{t, t^*}$
given $X_{t, t^*}=i^*$ and $X_t=i$ is governed by a parameter  $\theta^{(i,i^*)}$, and has density or probability mass function denoted $f^{*(i,i^*)}\left(\cdot; \theta^{(i,i^*)}\right)$ or sometimes simply denoted by $f^{*(i,i^*)}(\cdot)$.
Let $\Theta^{*(i)}=\left(\theta^{(i,1)}, \ldots, \theta^{(i,N^*)}\right)$, the fine-scale emission parameter vector corresponding to $X_t=i$.
%
\hfill\break
{\bf{NH:  will there be some parameters that are common across the fine-scale emission distributions? Do we want to separate them out?\\
Also - I used left ( and right ) above to make the parentheses a little bigger.  I think it looks better.  So we may want to do this throughout.
$f^{*(i,i^*)}\left(\cdot; \theta^{(i,i^*)}\right)$ versus
$f^{*(i,i^*)}(\cdot; \theta^{(i,i^*)})$.
}}
\hfill\break
%
The dependencies among the processes are defined as follows.
%
Given the coarse-scale states, $X_1,\ldots, X_T$, and the coarse scale emissions,  
$Y_1,\ldots, Y_T$, the $T$ fine-scale processes $(X_1^*, Y_1^*), \ldots, (X_T^*, Y_T^*)$, are independent HMMs.  
%%
\hfill\break
{\bf{I think that works for the dependencies - might need to add more. Does that look OK to you?}}
%%

@@

For more information on specific considerations for HHMMs such as incorporating covariates into the probability transition matrix, model selection and model checking, see Adam et al \citep{Adam:2019}.

@@

% do I need this subsections
\subsection{Decoding}

{\bf{I'm stopping here.  Perhaps you can take a stab at revisions here?}}

>>>>

%
We denote the conditional distribution of $Y_t$ given $Y_{t-1}= y_{t-1}$ and $ X_t=i$ as $f^{(i)}( \cdot | y_{t-1})$ or as
$f^{(i)}( \cdot | y_{t-1}; \theta^{(i)})$.
As an example, we could assume that this conditional distribution is normal with parameters $\theta^{(i)} = \{\mu^{(i)},\sigma^{(i)},\phi^{(i)}\}$ where:
%
\[
\mathbb{E}(Y_t|Y_{t-1} = y_{t-1},X_t=i) = \phi^{(i)} ~ y_{t-1} ~+ ~(1-\phi^{(i)})  ~\mu^{(i)}
\]
and
\[
\mathbb{V}(Y_t| Y_{t-1} =y_{t-1}, X_t = i) = (\sigma^{(i)})^2.
\]
%
The CarHMM therefore explicitly models auto-correlation into the emission distributions of the HMM while maintaining the structure needed to write the likelihood using the forward algorithm. In our normal model formulation, we have added only one additional parameter, $\phi^{(i)}$,  per possible hidden state.

The likelihood for the CarHMM can be easily calculated using the forward algorithm. As previously, let $\bf y$ be the vector of observed emissions. Then
\begin{equation}
\calL_{\text{CarHMM}}(y;\Theta,\Gamma) = \delta \prod_{t=2}^T \Gamma P(y_t|y_{t-1};\Theta) \mathbf{1}_N
\label{CarHMM_likelihood}
\end{equation}
where
%
$P(y_t|y_{t-1};\Theta)$ is an $N \times N$ diagonal matrix with $ii$th entry equal to $f^{(i)}(y_t|y_{t-1}; \theta^{(i)})$.
%

\subsection{The hierarchical hidden Markov model (HHMM)}

A \textit{hierarchical hidden Markov model} contains both a coarse-scale process and a fine-scale process. The coarse-scale process is a hidden Markov model as defined in the previous subsection, with $X_1, \ldots, X_T$ the unobserved Markov chain of states and $Y_1,\ldots, Y_T$ the corresponding observed responses. The possible states are $1,\ldots, N$.   
%
For each value of $t$, the state $X_t$ emits not only $Y_t$ but also a sequence of unobserved states, $X_t^* \equiv (X_{t,1}^*,\ldots, X_{t,T_t^*})$ and a sequence of observed emissions, $Y_t^* \equiv (Y_{t,1}^*,\ldots, Y_{t,T_t^*})$. These form our fine-scale process. The length of each fine-scale process, $T^*_t$, is a problem-specific tuning parameter that is determined by how the fine-scale observations are split. Conditional on the coarse-scale hidden state $X_t$, the joint fine-scale process, $X_t^*$, $Y_t^*$, is a hidden Markov model with parameters depending on the value of $X_t$.  Specifically, given that $X_t=i$, $X_t^*$ is a Markov chain with states $1,\ldots, N_t^*$. The distribution of $X_t^*$, conditional on $X_t=i$, is given by an $N^*_t \times N^*_t$ transition probability matrix $\Gamma^{*(i)}$ and initial probability, denoted by the $N_t^*$-vector $\delta^{*(i)}$, which we take equal to the stationary distribution of the chain. For simplicity, we take $N_t^* \equiv N^*$ although this is not necessary.
%
The distribution of $Y_{t, t^*}$ given $X_{t, t^*}=i^*$ and $X_t=i$ is governed by a parameter $\theta^{(i,i^*)}$, and has density or probability mass function denoted $f^{*(i,i^*)}\left(\cdot; \theta^{(i,i^*)}\right)$ or sometimes simply denoted by $f^{*(i,i^*)}(\cdot)$. Let $\Theta^{*(i)}=\left(\theta^{(i,1)}, \ldots, \theta^{(i,N^*)}\right)$, the fine-scale emission parameter vector corresponding to $X_t=i$.
%
The dependencies among the processes are defined as follows.
%
Given the coarse-scale states, $X_1,\ldots, X_T$, and the coarse scale emissions,  
$Y_1,\ldots, Y_T$, the $T$ fine-scale processes $(X_1^*, Y_1^*), \ldots, (X_T^*, Y_T^*)$, are independent HMMs. It is also possible to force certain parameters to be shared across different coarse or fine states. For example, in our killer whale case study we force certain fine-scale parameters to be shared across coarse-scale hidden states, i.e. $\theta^{(1,i*)} = \ldots = \theta^{(N,i*)}$.

@@

For more information on specific considerations for HHMMs such as incorporating covariates into the probability transition matrix, state decoding, model selection and model checking, see \citep{Adam:2019}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{General hierarchical structures withing HMMs}
\label{subsec:STFT}

One issue with all HMMs is that they assume Markovian dynamics when conditioned on the hidden state, i.e. that any observation $Y_t$ depends only on the behavioral state $X_t$ and $Y_{t-1}$ when conditioned on all previous time steps. However, there are many processes which violate this Markov property on very fine scales. Using the field of ecology as an example, swimming behaviour of marine mammals often exhibits periodic behavior since the animal repeatedly flukes (swims up and down) to propel itself forward. Work has been done in the past to model non-Markovian dynamics in the \textit{behavioural} process $X^*_t$ \citep{Langrock:2012}, but to our knowledge addressing non-Markovian dynamics within the observation process $Y^*_t$ is still a relatively unstudied area \textcolor{red}{(MAYBE?)}. With improvements in tagging technology allowing for data to be collected at very high frequencies, noisy and non-Markovian fine scale behavior is likely to persist.

To deal with fine-scale structures in the data which violate the Markov property, the fine-scale process $Y^*_t$ can be modeled with \textbf{any} parametric model which admits an easy-to compute likelihood. We remind the reader that the size of each chunk, $T^*_t$, is a problem-specific tuning parameter. It should be long enough to capture the fine-scale behavior within each chunk, but short enough to avoid over-smoothing of the data and to maintain high resolution in the hidden process $X^*$. The state-specific emission distribution $\calL_{\text{HMM}}$ from the HHMM likelihood is then replaced by the likelihood of the fine-scale model, $\calL_{\text{fine}}(\mathbf{y}_t;\Theta^{(i)})$. This gives the following likelihood function:
\[
\calL_{\text{coarse}}(y,y^*;\Theta,\Theta^*,\Gamma) = \delta P(\mathbf{y}_1;\Theta) \prod_{t=2}^T \Gamma P(y,y_t;\Theta,\Theta^*) \mathbf{1}_N
\]
where $P(\mathbf{y}_t;\Theta)$ is an $N \times N$ diagonal matrix with $ii$th entry corresponding to $X_t=i$ and equal to $f^(i)\left(\cdot\right)\calL_{\text{fine}}\left(y^*_t;\Theta^{*(i)}\right)$. This definition is straightforward to extend to the CarHMM on both the coarse- and fine- scale.

We describe one example of a fine-scale model which can simultaneously deal with non-Markovian, periodic behavior and reduce the dimension of the fine-scale process $y^*$.

\subsubsection{The short-time Fourier transform}

Suppose that the fine scale process $y^*_t$ exhibits significant period behavior which cannot be modeled with a Markov chain. One alternative is to use the Fourier transform of $y^*_t$: with a fixed window length $h$:
%
\begin{align*}
    STFT\{z_{(th+1):(th+h)}\}(k) := \hat{z}^{(k)}_{t} = \sum_{n = 1}^{h} z_{(th)+n}e^{-i \frac{2\pi k}{h} (n-1)} \\ k = 0, 1, \ldots, h-1; \quad t = 0,1, \ldots, \lfloor S/h \rfloor - 1
\end{align*}
%
where $i$ in the equation above refers the $\sqrt{-1}$. The STFT slides a moving window of length $h$ across the time series $\mathbf{z}$ and transforms the domain of each window from time to frequency. This allows the spectrum of $\mathbf{z}$ at time $s$ to be summarized by a $h$-dimensional vector of Fourier coefficients. We prevent any overlap in sliding windows to avoid serial dependence. $t$ indexes the \textit{windows} moving across $\mathbf{z}$ while $s$ indexes the \textit{observations} of $\mathbf{z}$.

If $\mathbf{z} \in \mathbb{R}^{S}$, then $\hat{\mathbf{z}} \in \mathbb{C}^{\lfloor S / h \rfloor \times h}$, which may be very large. Summary statistics can drastically reduce the dimension of $\hat{z}_s$. One possible example is to use the following:
%
$$y_t^{(1)} = \mathcal{R}\left(\hat{z}^{(0)}_t\right) \qquad y_t^{(2)} = \frac{1}{h}\sum_{k=1}^{\tilde{f}}|\hat{z}^{(k)}_s|^2$$
%
$y_t^{(1)}$ is equal to the average value of $\mathbf{z}$ within window $t$, and $y_t^{(2)}$ is equal to the squared 2-norm of the component of $\mathbf{z}$ within the window that can be attributed to frequencies between $1$ and $\tilde{f}$ periods per window length. Like $h$, the max frequency $\tilde{f}$ is a problem-specific tuning parameter. It should be selected such that $\tilde{f}$ periods per window length is the maximum frequency of $\mathbf{z}$ that makes biological sense. These summary statistics are just one possible choice to describe each window; other choices include the maximum and argmax of $\hat z_t$. Figure (\ref{fig:fourier_example}) visually shows the process of moving from $\mathbf{z}$ to $\mathbf{y}$.

\begin{figure}[ht]
	\centering
	\includegraphics[width=5in]{../Plots/fourier_transform.png}
	\caption{Visualization of transforming $\mathbf{z}_t$ into $\mathbf{y}_t$ using a sliding window and Fourier transform.}
	\label{fig:fourier_example}
\end{figure}

It is possible to accommodate for unequal time steps within each window by using the \textbf{non-uniform discrete Fourier transform (NDFT)}. We do not describe the details of this method in this work, but the generalization is straightforward. Refer to Bagchi et al \citep{Bagchi:1999} for details.

### 2020-07-08 Report/Sections/simulation.tex
Commit before changing STFT to Fourier Transform

>>>>

\subsection{Model Formulation}
\label{subsec:model_structure}

Hierarchical hidden Markov models can be used to jointly model simultaneous coarse-scale and fine-scale processes. However, as mentioned before, the fine-scale process $Y^*$ can often exhibit auto-correlation and intricate structure. Transforming $Y^*_t$ to $Z^*_t$ removes fine-scale periodic behavior, but $Z^*_t$ can still exhibit auto-correlation, especially in $Z_{t,s^*}^{*(1)}$. Therefore, we replace the fine-scale HMM within the hierarchical HMM with a CarHMM as shown in (fig \ref{fig:CarHHMM}).
%
\begin{figure}[ht]
	\centering
	\includegraphics[width=5in]{../Plots/CarHHMM.png}
	\caption{Graphical representation of a CarHHMM. The additional arrows representing auto-correlation between observations are shown in red for emphasis.}
	\label{fig:CarHHMM}
\end{figure}
%
Naturally, we refer to this model as the \textbf{CarHHMM}, and it can be implemented both with and without the short-time Fourier-transform (STFT). The likelihood of this model is still easy to calculate using the forward algorithm:
%
$$\calL_{\text{CarHHMM}}(y,z^*;\Theta,\Theta^*,\Gamma,\Gamma^*) = \delta P(y_1,z_1^*;\Theta,\Theta^*,\Gamma^*) \prod_{t=2}^T \Gamma P(y_t,z_t^*;\Theta,\Theta^*,\Gamma^*) \mathbf{1}_N$$
%
where:
%
\begin{align*}
P(y_t,z_t^*;\Theta,\Theta^*,\Gamma^*)  = \text{diag}\Big[&f^{(1)}(y_t)\calL_{\text{CarHMM}}\left(z_t^*;\Theta^{*(1)},\Gamma^{*(1)}\right), \ldots , \\
&f^{(N)}(y_t)\calL_{\text{CarHMM}}\left(z_t^*;\Theta^{*(N)},\Gamma^{*(N)}\right) \Big]
\end{align*}
